Synthesizing Quotient Lenses

### 2018-03-16 papers/QuotientLenses/main.tex
Tweaks

\begin{document}
\title{Synthesizing Quotient Lenses}
\begin{abstract}
{\em Quotient lenses} are bidirectional transformations whose correctness
laws are ``loosened'' by specified equivalence relations, allowing
inessential details in concrete data formats to be suppressed. 
For example, a programmer could use a quotient lens to define 
a transformation that ignores the order of fields in XML data, so
that two XML files with the same fields but in different orders would be
considered the same, allowing a single, simple program to handle them both. 
%
Building on a recently published algorithm for synthesizing plain bijective
lenses from high-level specifications, we show how to synthesize bijective
quotient lenses in three steps. First, we introduce {\em quotient regular
  expressions} (QREs), annotated regular expressions that conveniently mark
inessential aspects of string data formats; each QRE specifies,
simulteneously, a regular language and an equivalence relation on it.
Second, we introduce {\em QRE lenses}, i.e., lenses mapping between QREs.
Our key technical result is a proof that every QRE lens can be transformed
into a functionally equivalent lens that canonizes source and target data only
once at the ``edges'' then uses a bijective lens to map between the respective
canonical elements; no internal canonization occurs in a lens in this normal
form. Third, we leverage this normalization theorem to {\em synthesize} QRE
lenses from a pair of QREs and example input-output pairs, reusing earlier work
on synthesizing plain bijective lenses. We have implemented QREs and QRE lens
synthesis as an extension to the bidirectional programming language Boomerang.
We evaluate the effectiveness of our approach by synthesizing QRE lenses
between various real-world data formats in the Optician benchmark suite.
\end{abstract}

>>>>

\begin{document}
\title{Synthesizing Quotient Lenses}
\begin{abstract}
{\em Quotient lenses} are bidirectional transformations whose correctness
laws are ``loosened'' by specified equivalence relations, allowing
inessential details in concrete data formats to be suppressed. 
For example, a programmer could use a quotient lens to define 
a transformation that ignores the order of fields in XML data, so
that two XML files with the same fields but in different orders would be
considered the same, allowing a single, simple program to handle them both. 
%
Building on a recently published algorithm for synthesizing plain bijective
lenses from high-level specifications, we show how to synthesize bijective
quotient lenses in three steps. First, we introduce {\em quotient regular
  expressions} (QREs), annotated regular expressions that conveniently mark
inessential aspects of string data formats; each QRE specifies,
simulteneously, a regular language and an equivalence relation on it.
Second, we introduce {\em QRE lenses}, i.e., lenses mapping between QREs.
Our key technical result is a proof that every QRE lens can be transformed
into a functionally equivalent lens that canonizes source and target data just
at the ``edges'' and that uses a bijective lens to map between the respective
canonical elements; no internal canonization occurs in a lens in this normal
form. Third, we leverage this normalization theorem to {\em synthesize} QRE
lenses from a pair of QREs and example input-output pairs, reusing earlier work
on synthesizing plain bijective lenses. We have implemented QREs and QRE lens
synthesis as an extension to the bidirectional programming language Boomerang.
We evaluate the effectiveness of our approach by synthesizing QRE lenses
between various real-world data formats in the Optician benchmark suite.
\end{abstract}
